Data Quality - Data Analysis

### 2025-03-09 study2/analysis/2_analysis.qmd
update

df <- read.csv("../data/data_participants.csv") |> 
  filter(!is.na(MINT_Awareness_ExAc_35))
```

@@

items <- select(df, starts_with("MINT_"))
names(items) <- str_remove(names(items), "MINT_")

colors <- c(MINT="#00BCD4", MINTi="#00838F", IAS="#F44336", MAIA="#FF9800", MAIAi="#EF6C00", BPQ="#795548")
```

@@

@@

```{r}
compare_models <- function(outcome="TAS_DIF") {

@@

  test <- test_performance(models)

@@

  # Full Questionnaires
  m_mint <- lm(as.formula(paste0(outcome, " ~ .")),
              data = select(df, all_of(outcome), starts_with("MINT"), -matches("[[:digit:]]$"))) |> 
    step(direction = "both", trace=0)
  m_minti <- lm(as.formula(paste0(outcome, " ~ .")),
              data = cbind(minti, df[outcome])) |> 
    step(direction = "both", trace=0)
  m_ias <- lm(as.formula(paste0(outcome, " ~ IAS")),
              data = df)
  m_maia <- lm(as.formula(paste0(outcome, " ~ .")),
                data = select(df, all_of(outcome), starts_with("MAIA"))) |> 
    step(direction = "both", trace=0)
  m_maiai <- lm(as.formula(paste0(outcome, " ~ .")),
                data = select(df, all_of(outcome), 
                              starts_with("MAIA"), 
                              -contains("NotWorrying"),
                              -contains("NotDistracting"))) |> 
    step(direction = "both", trace=0)
  m_bpq <- lm(as.formula(paste0(outcome, " ~ BPQ")),
                data = df)

  test_all <- test_performance(m_mint, m_minti, m_ias, m_maia, m_maiai, m_bpq)
  test_all$Predictors <- c(
    paste0(cleanlabels(find_predictors(m_mint)$conditional), collapse = ", "),
    paste0(cleanlabels(find_predictors(m_minti)$conditional), collapse = ", "),
    "", 
    paste0(cleanlabels(find_predictors(m_maia)$conditional), collapse = ", "),
    paste0(cleanlabels(find_predictors(m_maiai)$conditional), collapse = ", "),
    "")

  list(test=test,  test_all=test_all, p=p)
}

rez1 <- compare_models("TAS_DIF")
rez1$test
rez1$p
rez1$test_all

rez2 <- compare_models("TAS_DDF")
rez2$test
rez2$p
rez2$test_all

rez3 <- compare_models("TAS_EOT")
rez3$test
rez3$p
rez3$test_all
```

@@

```{r}
rez <- compare_models("LifeSatisfaction")
rez$test
rez$p
rez$test_all

rez <- compare_models("PHQ4_Anxiety")
rez$test
rez$p
rez$test_all

rez <- compare_models("PHQ4_Depression")
rez$test
rez$p
rez$test_all
```

@@

```{r}
rez <- compare_models("ERS_Arousal")
rez$test
rez$p
rez$test_all

rez <- compare_models("ERS_Sensitivity")
rez$test
rez$p
rez$test_all

rez <- compare_models("ERS_Persistence")
rez$test
rez$p
rez$test_all
```

@@

```{r}
compare_predictive <- function(outcome="Disorders_Psychiatric_Mood") {
  m_mint <- glm(as.formula(paste0(outcome, " ~ .")),
              data = select(df, all_of(outcome), starts_with("MINT"), -matches("[[:digit:]]$")),
              family = "binomial") |> 
  step(direction = "both", trace=0)

@@

@@

  test <- test_performance(m_mint, m_minti, m_ias, m_maia, m_maiai, m_bpq)

@@

  p <- rbind(
    mutate(as.data.frame(performance_roc(m_mint)), Predictor = "MINT"),
    mutate(as.data.frame(performance_roc(m_minti)), Predictor = "MINTi"),
    mutate(as.data.frame(performance_roc(m_ias)), Predictor = "IAS"),
    mutate(as.data.frame(performance_roc(m_maia)), Predictor = "MAIA"),
    mutate(as.data.frame(performance_roc(m_maiai)), Predictor = "MAIAi"),
    mutate(as.data.frame(performance_roc(m_bpq)), Predictor = "BPQ")
  ) |> 
      ggplot(aes(x=Specificity)) +
      geom_abline(intercept=0, slope=1, color="gray") +
      geom_line(aes(y=Sensitivity, color=Predictor), linewidth=2, alpha=0.9) +
      geom_label(data=data.frame(
        label = c(get_auc(m_mint), get_auc(m_minti), get_auc(m_ias), get_auc(m_maia), get_auc(m_maiai), get_auc(m_bpq)),
        Predictor = c("MINT", "MINTi", "IAS", "MAIA", "MAIAi", "BPQ"),
        x = 0.25,
        y = c(0.65, 0.60, 0.55, 0.45, 0.35, 0.20)), 
        aes(x=x, y=y, label=label), color=colors) +
      labs(x= "1 - Specificity (False Positive Rate)", y="Sensitivity (True Positive Rate)",
           color="Questionnaire Version") +
      scale_color_manual(values=colors) +
      scale_x_continuous(labels=scales::percent) +
      scale_y_continuous(labels=scales::percent) +
      theme_minimal()

  list(m_mint=m_mint, m_ias=m_ias, m_maia=m_maia, m_maiai=m_maiai, test=test, p=p)
}
```

```{r}
rez <- compare_predictive("Disorders_Psychiatric_Mood")
rez$p
```

```{r}
rez <- compare_predictive("Disorders_Psychiatric_MoodTreatment")
rez$p
```

@@

```{r}
model <- glm(as.formula(paste0("Wearables_Number ~ .")),
              data = select(df, Wearables_Number, starts_with("MINT"), -matches("[[:digit:]]")),
              family = "poisson") |> 
  step(direction = "both", trace=0)

parameters::parameters(model)

model <- glm(as.formula(paste0("Wearables_Heart ~ .")),
              data = df |>
              mutate(Wearables_Heart = Wearables_Heart != "Not owning") |> 
              select(Wearables_Heart, starts_with("MINT"), -matches("[[:digit:]]")),
              family = "binomial") |> 
  step(direction = "both", trace=0, k = log(nrow(df)))
summary(model)

model <- lm(as.formula(paste0("Wearables_HeartImportance ~ .")),
              data = df |>
              filter(!is.na(Wearables_HeartImportance)) |> 
              select(Wearables_HeartImportance, starts_with("MINT"), -matches("[[:digit:]]"))) |> 
  step(direction = "both", trace=0, k = log(nrow(filter(df, !is.na(Wearables_HeartImportance)))))

parameters::parameters(model)
```

>>>>

df <- read.csv("../data/data_participants.csv")
```

@@

items <- filter(df, !is.na(MINT_Awareness_ExAc_35)) 
names(items) <- str_remove(names(items), "MINT_")

colors <- c(MINT="#00BCD4", MINTi="#00838F", IAS="#F44336", MAIA="#FF9800", MAIAi="#EF6C00", MAIAmin="#E65100", BPQ="#795548")
```

@@

Most correlated dims:
- CaCo x8
- Urin x4
- Rela x4
- Derm x3
- Sati
- Olfa
- StaS
- ExaC
- Resp

@@

```{r}
table_models <- function(models) {
  test <- test_performance(models)
  test$Model <- NULL
  test$Omega2 <- NULL
  test$p_Omega2 <- NULL
  test$R2 <- compare_performance(models, metrics="R2")$R2

  preds <- c()
  for(m in models){
    x <- names(sort(abs(coef(m)[-1]), decreasing=TRUE))
    preds <- c(preds, paste0(cleanlabels(x), collapse = ", "))
  }
  test$Predictors <- preds
  format(test, zap_small = TRUE)
}

compare_models <- function(outcome="TAS_DIF") {

@@

  test <- table_models(models)
  test$Predictors <- NULL

@@

  # Full Questionnaires
  m_mint <- lm(as.formula(paste0(outcome, " ~ .")),
              data = select(df, all_of(outcome), starts_with("MINT"), -matches("[[:digit:]]$"))) |> 
    step(direction = "both", trace=0)
  m_minti <- lm(as.formula(paste0(outcome, " ~ .")),
              data = cbind(minti, df[outcome])) |> 
    step(direction = "both", trace=0)
  m_ias <- lm(as.formula(paste0(outcome, " ~ IAS")),
              data = df)
  m_maia <- lm(as.formula(paste0(outcome, " ~ .")),
                data = select(df, all_of(outcome), starts_with("MAIA"))) |> 
    step(direction = "both", trace=0)
  m_maiai <- lm(as.formula(paste0(outcome, " ~ .")),
                data = select(df, all_of(outcome), 
                              starts_with("MAIA"), 
                              -contains("NotWorrying"),
                              -contains("NotDistracting"))) |> 
    step(direction = "both", trace=0)
  m_maiamin <- lm(as.formula(paste0(outcome, " ~ .")),
                data = select(df, all_of(outcome),
                              MAIA_AttentionRegulation, # remove?
                              MAIA_SelfRegulation, # remove?
                              # MAIA_Trusting, # remove?
                              MAIA_BodyListening, 
                              MAIA_EmotionalAwareness,
                              MAIA_Noticing)) |> 
    step(direction = "both", trace=0)
  m_bpq <- lm(as.formula(paste0(outcome, " ~ BPQ")),
                data = df)

  models_full <- list(m_mint=m_mint, m_minti=m_minti, m_maia=m_maia, m_maiai=m_maiai, m_maiamin=m_maiamin, m_ias=m_ias, m_bpq=m_bpq)
  test_all <- table_models(models_full)

  list(test=test,  test_all=test_all, p=p, models_full=models_full)
}

rez1 <- compare_models("TAS_DIF")
display(rez1$test, title="DIF")
rez1$p
display(rez1$test_all, title="DIF")

rez2 <- compare_models("TAS_DDF")
display(rez2$test, title="DDF")
rez2$p
display(rez2$test_all, title="DDF")

rez3 <- compare_models("TAS_EOT")
display(rez3$test, title="EOT")
rez3$p
display(rez3$test_all, title="EOT")
```

```{r}
make_cor(select(df, starts_with("TAS_")),
         select(df, starts_with("MINT"), -matches("[[:digit:]]"), 
                starts_with("MAIA"), IAS, BPQ))
```

CaCo and Sati.

@@

```{r}
rez <- compare_models("LifeSatisfaction")
display(rez$test, title="Life Satisfaction") 
rez$p
display(rez$test_all, title="Life Satisfaction")

rez <- compare_models("PHQ4_Anxiety")
display(rez$test, title="Anxiety") 
rez$p
display(rez$test_all, title="Anxiety")

rez <- compare_models("PHQ4_Depression")
display(rez$test, title="Depression") 
rez$p
display(rez$test_all, title="Depression")
```

```{r}
make_cor(select(df, LifeSatisfaction, starts_with("PHQ4_")),
         select(df, starts_with("MINT"), -matches("[[:digit:]]"), 
                starts_with("MAIA"), IAS, BPQ))
```

- CaCo and RelA

@@

```{r}
rez <- compare_models("ERS_Arousal")
display(rez$test, title="Arousal") 
rez$p
display(rez$test_all, title="Arousal")

rez <- compare_models("ERS_Sensitivity")
display(rez$test, title="Sensitivity") 
rez$p
display(rez$test_all, title="Sensitivity")

rez <- compare_models("ERS_Persistence")
display(rez$test, title="Persistence") 
rez$p
display(rez$test_all, title="Persistence")
```

```{r}
make_cor(select(df, starts_with("ERS_")),
         select(df, starts_with("MINT"), -matches("[[:digit:]]"), 
                starts_with("MAIA"), IAS, BPQ))
```

- CaCo

@@

```{r}
compare_predictive <- function(outcome="Disorders_Psychiatric_Mood", y = c(0.65, 0.60, 0.50, 0.45, 0.40, 0.30, 0.20)) {
  m_mint <- glm(as.formula(paste0(outcome, " ~ .")),
              data = select(df, all_of(outcome), starts_with("MINT"), -matches("[[:digit:]]$")),
              family = "binomial") |> 
  step(direction = "both", trace=0)

@@

  m_maiamin <- glm(as.formula(paste0(outcome, " ~ .")),
                data = select(df, all_of(outcome),
                              MAIA_AttentionRegulation, # remove?
                              MAIA_SelfRegulation, # remove?
                              # MAIA_Trusting, # remove?
                              MAIA_BodyListening, 
                              MAIA_EmotionalAwareness,
                              MAIA_Noticing),
                family = "binomial") |> 
    step(direction = "both", trace=0)

@@

  test <- table_models(list(m_mint=m_mint, m_minti=m_minti, m_maia=m_maia, m_maiai=m_maiai, m_maiamin=m_maiamin, m_ias=m_ias, m_bpq=m_bpq))

@@

  p <- rbind(
    mutate(as.data.frame(performance_roc(m_mint)), Predictor = "MINT"),
    mutate(as.data.frame(performance_roc(m_minti)), Predictor = "MINTi"),
    mutate(as.data.frame(performance_roc(m_maia)), Predictor = "MAIA"),
    mutate(as.data.frame(performance_roc(m_maiai)), Predictor = "MAIAi"),
    mutate(as.data.frame(performance_roc(m_maiamin)), Predictor = "MAIAmin"),
    mutate(as.data.frame(performance_roc(m_ias)), Predictor = "IAS"),
    mutate(as.data.frame(performance_roc(m_bpq)), Predictor = "BPQ")
  ) |> 
    mutate(Predictor = fct_relevel(Predictor, "MINT", "MINTi", "MAIA", "MAIAi", "MAIAmin", "IAS", "BPQ")) |> 
      ggplot(aes(x=Specificity)) +
      geom_abline(intercept=0, slope=1, color="gray") +
      geom_line(aes(y=Sensitivity, color=Predictor), linewidth=2, alpha=0.9) +
      geom_label(data=data.frame(
        label = c(get_auc(m_mint), get_auc(m_minti), get_auc(m_maia), get_auc(m_maiai), get_auc(m_maiamin), get_auc(m_ias), get_auc(m_bpq)),
        Predictor = c("MINT", "MINTi", "MAIA", "MAIAi", "MAIAmin", "IAS", "BPQ"),
        x = 0.9,
        y = y), 
        aes(x=x, y=y, label=label), color=colors[c("MINT", "MINTi", "MAIA", "MAIAi", "MAIAmin", "IAS", "BPQ")]) +
      labs(x= "1 - Specificity (False Positive Rate)", y="Sensitivity (True Positive Rate)",
           color="Questionnaire Version") +
      scale_color_manual(values=colors) +
      scale_x_continuous(labels=scales::percent) +
      scale_y_continuous(labels=scales::percent) + 
      theme_minimal()

  list(m_mint=m_mint, m_ias=m_ias, m_maia=m_maia, m_maiai=m_maiai, m_maiamin=m_maiamin, test=test, p=p)
}
```

```{r}
rez <- compare_predictive("Disorders_Psychiatric_Mood")
rez$p
display(rez$test, title="Mood")
```

```{r}
rez <- compare_predictive("Disorders_Psychiatric_MoodTreatment")
rez$p
display(rez$test, title="Mood (with Treatment)")
```

@@

```{r}
model <- glm(as.formula(paste0("Wearables_Number ~ .")),
              data = select(df, Wearables_Number, starts_with("MINT"), -matches("[[:digit:]]")),
              family = "poisson") |> 
  step(direction = "both", trace=0)
parameters::parameters(model)

# Heart
m_owns <- glm(as.formula(paste0("Wearables_Heart ~ .")),
              data = df |>
              mutate(Wearables_Heart = Wearables_Heart != "Not owning") |> 
              select(Wearables_Heart, starts_with("MINT"), -matches("[[:digit:]]")),
              family = "binomial") |> 
  step(direction = "both", trace=0, k = log(nrow(df)))
parameters::parameters(m_owns)

model <- lm(as.formula(paste0("Wearables_HeartImportance ~ .")),
              data = df |>
              filter(!is.na(Wearables_HeartImportance)) |> 
              select(Wearables_HeartImportance, starts_with("MINT"), -matches("[[:digit:]]"))) |> 
  step(direction = "both", trace=0, k = log(nrow(filter(df, !is.na(Wearables_HeartImportance)))))
parameters::parameters(model)

lm(Wearables_HeartImportance ~ MINT_Interoception_Card, data=df) |> 
  estimate_relation() |> 
  plot()
```

### 2025-03-09 study2sona/analysis/analysis.qmd
update

@@

```{r}
dfchecks <- df |>
  dplyr::mutate(
    # "I can always accurately answer to the extreme left on this question to show that I am reading it"
    A1 = ifelse(MINT_AttentionCheck_1 == 0, 0, 1),
    # "I notice that I am being asked to respond all the way to the right"
    A2 = ifelse(MAIA_AttentionCheck_1 == 6, 0, 1),
    # "I can always accurately choose the lowest option"
    A3 = ifelse(IAS_AttentionCheck_1 == 1, 0, 1),
    # "Respond all the way to the right."
    A4 = ifelse(BodyAwareness_AttentionCheck_1 == 5, 0, 1),
    # "I am able to respond all the way to the left"
    A5 = ifelse(TAS_AttentionCheck_1 == 1, 0, 1),
    # "On the whole, I know I must press the highest option"
    A6 = ifelse(PI18_AttentionCheck_1 == 5, 0, 1),
    # "I feel that to show I'm being attentive I will press the lowest option"
    A7 = ifelse(CEFSA_AttentionCheck_1 == 0, 0, 1),
    .keep = "none"
  ) 
dfchecks$TotalFailed <- rowSums(dfchecks)
df$TotalFailed <- dfchecks$TotalFailed

@@

```{r}
df |> 
  mutate(Failed = ifelse(TotalFailed >= 1, 1, 0)) |> 
  summarize(Failed = sum(Failed), N = n(), p = Failed / N, .by="Condition") |> 
  gt::gt()
```

```{r}
m1 <- glm(TotalFailed ~ Condition, data=df, family="poisson")
m2 <- glmmTMB::glmmTMB(TotalFailed ~ Condition,
                      data=df,
                      ziformula=~1,
                      family="poisson")
m3 <- glmmTMB::glmmTMB(TotalFailed ~ Condition,
                      data=df,
                      ziformula=~1,
                      family=glmmTMB::truncated_nbinom2)

@@

modelbased::estimate_means(m1, by="Condition", backend="emmeans") |> 
  plot()
```

>>>>

```{r}

```

@@

```{r}
dfchecks <- df |>
  dplyr::mutate(
    # "I can always accurately answer to the extreme left on this question to show that I am reading it"
    A1 = ifelse(MINT_AttentionCheck_1 == 0, 0, 1),
    # "I notice that I am being asked to respond all the way to the right"
    A2 = ifelse(MAIA_AttentionCheck_1 == 6, 0, 1),
    # "I can always accurately choose the lowest option"
    A3 = ifelse(IAS_AttentionCheck_1 == 1, 0, 1),
    # "Respond all the way to the right."
    A4 = ifelse(BodyAwareness_AttentionCheck_1 == 5, 0, 1),
    # "I am able to respond all the way to the left"
    A5 = ifelse(TAS_AttentionCheck_1 == 1, 0, 1),
    # "On the whole, I know I must press the highest option"
    A6 = ifelse(PI18_AttentionCheck_1 == 5, 0, 1),
    # "I feel that to show I'm being attentive I will press the lowest option"
    A7 = ifelse(CEFSA_AttentionCheck_1 == 0, 0, 1),
    .keep = "none"
  ) 
dfchecks$TotalFailed <- rowSums(dfchecks)
df$TotalFailed <- dfchecks$TotalFailed
df$Failed <- ifelse(dfchecks$TotalFailed >= 1, 1, 0)

@@

```{r}
df |> 
  summarize(Failed = sum(Failed), N = n(), p = Failed / N, .by="Condition") |> 
  gt::gt()
```

### Failed at least one

```{r}
m1 <- glm(Failed ~ Prolific, 
          data=mutate(df, Prolific = ifelse(Condition == "Prolific", "Prolific", "SONA")), 
          family="binomial")
# m1 <- brms::brm(Failed ~ Condition, data=df, family="poisson", backend="cmdstanr")

modelbased::estimate_means(m1, by="Prolific", keep_iterations=10) |> 
  plot()

modelbased::estimate_contrasts(m1)
```

```{r}
m1 <- glm(Failed ~ Condition, data=df, family="binomial")
# m1 <- brms::brm(Failed ~ Condition, data=df, family="poisson", backend="cmdstanr")

modelbased::estimate_means(m1, by="Condition", keep_iterations=10) |> 
  plot()

modelbased::estimate_contrasts(m1)
```

### Number

```{r}
m1 <- glm(TotalFailed ~ Condition, data=df, family="poisson")
m2 <- glmmTMB::glmmTMB(TotalFailed ~ Condition,
                      data=df,
                      ziformula=~Condition,
                      family="poisson")
m3 <- glmmTMB::glmmTMB(TotalFailed ~ Condition,
                      data=df,
                      ziformula=~Condition,
                      family=glmmTMB::truncated_nbinom2)

@@

modelbased::estimate_means(m1, by="Condition") |> 
  plot()
```
